Modelling the gas kinematics of an atypical Lyman alpha emitting compact dwarf galaxy

### 2017-06-21 paper/paper.tex
new abstract

\begin{abstract}
  Star-forming Compact Dwarf Galaxies (CDGs) resemble the expected
  pristine conditions of the first galaxies in the Universe and
  are the best systems to test models on primordial
galaxy formation and evolution.    
Here we report on one of such CDGs, \tol, which presents
a broad, single peaked, highly symmetric \lya\ line emission that had
evaded theoretical interpretation so far.  
In this paper we explain these features by two different physically
motivated kinematic models: 
an interstellar medium composed by outflowing clumps with 
random motions and an homogeneous gaseous sphere undergoing solid body
rotation.
It is the first time that an observed \lya\ spectrum can be explained
assuming either of these kinematic conditions.
We find that both models independently require high velocities
(either a clump velocity dispersion of \sigmaclump\ with outflows of
\inftyclump\ or a bulk rotation of $348^{+75}_{-48}$ km s$^{-1}$)
consistent with a dynamical mass of at 
least a $10^{10}$ M$_{\odot}$, ten times larger than its estimated
baryonic mass.     
This work shows the importance of considering multiphase
kinematics and rotation among the possible conditions shaping the
\lya\ spectra of the first galaxies.
Additionally, if future kinematic maps of tol confirm the velocities
found by either of our models, it would completeley justify the need for
detailed studies to understand the reasons behind a such a high
dynamical mass for a compact dwarf galaxy.
\end{abstract}

@@

\tol\ shows one of the most atypical \lya\ spectra observed so far.
Its flux at the line's center is high compared to other LAEs at low
redshift and the wide symmetrical peak is virtually absent from other
LAEs at low and high redshift \citep{2012ApJ...751...29Y,LARS,Erb14,Trainor16}. Simple
shell models fail to reproduce such a spectrum
\cite{2015A&A...578A...7V}.  
In the previous sections we show how these characteristics can be
explained by rather extreme kinematic characteristics.

A first perspective to put in context \tol's kinematic is comes the
observed scaling relations for elliptic galaxies.
Assuming that \tol\ follows the fundamental plane relationship
between its mean surface brightness $I_e=L_V/2\pi R_{e}^2$, the
projected half-light radius $R_e$ and the velocity dispersion
$\sigma$, described by $\log I_e=1.6 \log\sigma - 1.21\log R_e + 0.55$ \citep{2009ApJ...698.1590G},
the expected velocity dispersion should be on the order of $5\pm 1$
\kms, which is a factor of $\sim\ 10$ - $60$ lower than the   
results from the multiphase and rotation models, respectively.

However, \tol's structural properties, seem to be closer to observed dwarf
elliptical galaxies \citep{2003AJ....126.1794G}. 
Galaxies such as IC 3344 and NGC 4482 have similar absolute $M_V$
magnitudes  and size as \tol. 
These galaxies have velocity dispersions in the range $30$-$45$ \kms. 
The rotational velocities of those systems is lower than $1$\kms. 
This gives support to the idea that the multiphase model might be
closer to be the correct model describing \tol, as it provides at
least kinematic consistency with other dwarf galaxy populations.

>>>>

\begin{abstract}
  Star-forming Compact Dwarf Galaxies (CDGs) resemble the expected
  pristine conditions of the first galaxies in the Universe and
  are the best systems to test models on primordial
galaxy formation and evolution.    
Here we report on one of such CDGs, \tol, which presents
a broad, single peaked, highly symmetric \lya\ line emission that had
evaded theoretical interpretation so far.  
In this paper we explain these features by two different physically
motivated kinematic models: 
an interstellar medium composed by outflowing clumps with 
random motions and an homogeneous gaseous sphere undergoing solid body
rotation.
It is the first time that an observed \lya\ spectrum can be explained
assuming either of these kinematic conditions.
We find that both models independently require high velocities, either
a clump velocity dispersion of \sigmaclump\ with outflows of
\inftyclump\ or a bulk rotation of $348^{+75}_{-48}$ km s$^{-1}$. 
The velocity dispersion in the multiphase model is consistent with a dynamical
mass of $2\times 10^{9}$ M$_{\odot}$, ten times its baryonic mass,
which can be explained if \tol\ is hosted by a dark matter halo of at least
$10^{11}$ M$_{\odot}$. 
These conditions are similar to Dwarf Elliptical Galaxies.
This work shows the importance of considering multiphase
and rotation kinematics among the possible conditions shaping the
\lya\ spectra.
If future kinematic maps of \tol\ confirm the velocities found by the
multiphase model, it will be the basis to study new processes that
might be at play during priomordial galaxy evolution.
\end{abstract}

@@

\tol\ shows one of the most atypical \lya\ spectra observed so far.
Its flux at the line's center is high compared to other LAEs at low
redshift and the wide symmetrical peak is virtually absent from other
LAEs at low and high redshift \citep{2012ApJ...751...29Y,LARS,Erb14,Trainor16}. Simple
shell models fail to reproduce such a spectrum
\cite{2015A&A...578A...7V}.  
In the previous sections we show how these characteristics can be
explained by two different kinematic models. Which model should we prefer?

\tol's structural properties seem to be similar to observed dwarf
elliptical galaxies \citep{2003AJ....126.1794G}. 
Galaxies such as IC 3344 and NGC 4482 have similar absolute $M_V$
magnitudes and size as \tol. 
These galaxies have velocity dispersions in the range $30$-$45$ \kms. 
The rotational velocities of those systems is lower than $1$\kms. 
This supports the idea that the multiphase model the correct model
describing \tol, as it provides at least some consistency with
kinematic properties of other dwarf elliptical galaxies.
